---
title: "Interactive Maps"
author: "Parker Max"
format: html
---

**INTERACTIVE MAP #1: PCE by US STATE**
```{r}
library(leaflet)
library(sf)
library(leaflet) 
library(htmltools)
library(glue)
library(dplyr) #reading in all the needed libraries
PCE2 <- read.csv("spending.csv") #Reads the 2023 PCE Data

states <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson")
# Merge PCE data with states
states <- states |> 
  left_join(PCE2, by = c("name" = "State"))

# Create bins for PCE values
bins <- c(0, 200000, 700000, 1200000, 1700000, 2200000, Inf)
pal <- colorBin("YlOrRd", domain = states$X2023, bins = bins)

states <- states |> # Create labels for hover tooltips
  mutate(labels = glue("<strong>{name}</strong><br/>PCE: ${X2023} Million"))

labels <- lapply(states$labels, HTML)

leaflet(states) |> # Generate interactive map
  setView(-96, 37.8, 4) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal(X2023),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) |>
  addLegend(pal = pal, values = ~`X2023`, opacity = 0.7, title = "2023 PCE in Millions of Dollars",
    position = "bottomright")
```
This is an interactive leaflet map. It shows Personal Consumption Expenditure (PCE) data by state for 2023 in millions of dollars. The states are color coded using a gradient, where lighter colors represents lower PCE values and darker colors indicates higher values. Hovering over the states shows the data from the Bureau of Economic Analysis shows that Texas ($2,000M+), California ($2,500M+), and Florida ($1,300M+) have the highest consumer spending. In contrast, states in the Midwest and Mountain West generally have lower PCE, with many below $500M. This shows the regional differences in economic activity and consumer spending across the U.S.


**INTERACTIVE MAP #2: WHICH STATES HAVE AT LEAST ONE NATIONAL PARK?**


```{r}
nps_parks <- read_csv("nps_parks.csv")
nps_parks$state <- tolower(nps_parks$state) #when I manually entered the data, the states are lower case.
us_states <- map_data("state")
map2 <- us_states |> 
  left_join(nps_parks, by = c("region" = "state")) #joins the parks data with the mapping data
ggplot(map2, aes(long, lat, group = group, fill = factor(park))) + #treated as a categorical (factor) variable rather than a continuous one
  geom_polygon(color = "black") +
  scale_fill_manual(
    values = c("1" = "darkgreen", "0" = "chocolate4"), #the data is 0 = No, 1 = yes, there is a park in this state
    name = "Has a National Park",
    labels = c("No", "Yes")
  ) +
  theme_void() +
  labs(title = "US States with a National Park",
       caption = "Data Source: National Park Service & Wikipedia (Manual Entry)")



```
This choropleth map of the United States showing which states have at least one national park. States with at least one national park are shaded in dark green, while those without any are shaded in brown (the two colors of the National Parks Service!). The maps shows that the majority of US States has at least one national park in their state. The data on whether or not the states have a national park comes from the National Park Service and Wikipedia, then manually entered into a csv file. 



```{r}
states <- read_sf("https://rstudio.github.io/leaflet/json/us-states.geojson")
nps_parks_int <- read.csv("nps_parks.csv") #creates the nps_parks_int for interactive parks
# Merge nps data with states
states <- states |> 
  left_join(nps_parks_int, by = c("name" = "state"))

# Create bins for PCE values
pal <- colorFactor(
  palette = c("chocolate4", "darkgreen"), # No = brown, Yes = green, color scale, these are the colors of the NPS!
  domain = states$park
)


states <- states |> 
  mutate(labels = ifelse(park == 1, 
                         glue("<strong>{name}</strong><br/>Has at least one National Park"), 
                         glue("<strong>{name}</strong><br/>No National Park"))) #Creating the message that appears when I hover over a state

labels <- lapply(states$labels, HTML) 

leaflet(states) |> #creating the leaflet map
  setView(-96, 37.8, 4) |>
  addTiles() |>
  addPolygons(
    fillColor = ~pal(park),
    weight = 2,
    opacity = 1,
    color = "white",
    dashArray = "3",
    fillOpacity = 0.7,
    highlightOptions = highlightOptions(
      weight = 5,
      color = "#666",
      dashArray = "",
      fillOpacity = 0.7,
      bringToFront = TRUE),
    label = labels,
    labelOptions = labelOptions(
      style = list("font-weight" = "normal", padding = "3px 8px"),
      textsize = "15px",
      direction = "auto")) |>
  addLegend(
    pal = pal, 
    values = ~park, 
    labels = c("No", "Yes"), #labels 
    opacity = 0.7, 
    title = "Has a National Park?",
    position = "bottomright")

```
This an interactive leaflet map of the United States showing which states have at least one national park. States with at least one national park are shaded in dark green, while those without any are shaded in brown. From looking at the map and hovering over each state, it shows that the majority of US States has at least one national park in their state. The data on whether or not the states have a national park comes from the National Park Service and Wikipedia, then manually entered into a csv file. 

