---
title: "Mini Project Static Maps"
author: "Parker Max"
format: pdf
---

title: "Mini Project 1: Parker Max"
sidebar: false
format:
  pdf: default
editor_options:  
  chunk_output_type: console
---
---
```{r}
#| message: false
#| warning: false

# Initial packages required (we'll be adding more)
library(tidyverse)
library(mdsr)      # package associated with our MDSR book
```


**STATIC MAP #1: PCE DATA by US STATE**

```{r}
#| message: false
#| warning: false
library(maps) #loads the maps library
us_states <- map_data("state") #creates our US state Map
head(us_states)

PCE <- read.csv("spending.csv") #Reads the 2023 PCE Data
PCE$State <- tolower(PCE$State) #Converts the state names to be lower case to be consistent with our us_states map

map1 <- us_states |>
  left_join(PCE, by = c("region" = "State")) |> #joins the PCE data with the map
mutate(PCE_category = cut(X2023, 
    breaks = c(0, 200000, 700000, 1200000, 1700000, 2200000, Inf), 
    labels = c("< $200K", "$200k-$500k", "$700k-$1.2M", "$1.2M-$1.7M", "$1.7M-$2.2M", ">$2.5M")))
  
```


```{r}
static_map1 <- ggplot(map1,(aes(x = long, y = lat, group = group, fill = PCE_category))) + #creating my first static map
  geom_polygon(color = "white") +
  scale_fill_brewer( name = "PCE (Millions of USD)") +
  theme_void() +
  labs(x = "", y = "", title = "Personal Consumption Expenditure by US State in 2023 (USD)", subtitle = "Data Source: Bureau of Economic Analysis - https://www.bea.gov/data/consumer-spending/state")
print(static_map1)
```

This choropleth map of the United States shows Personal Consumption Expenditure (PCE) data by state for 2023 in millions of dollars. The states are color coded using a gradient, where lighter colors represents lower PCE values and darker blue indicates higher values. The data from the Bureau of Economic Analysis shows that Texas ($2,000M+), California ($2,500M+), and Florida ($1,500M+) have the highest consumer spending. In contrast, states in the Midwest and Mountain West generally have lower PCE, with many below $500M. This shows the regional differences in economic activity and consumer spending across the U.S. 

**STATIC MAP #2: WHICH STATES HAVE AT LEAST ONE NATIONAL PARK?**
```{r, echo = TRUE}
nps_parks <- read_csv("nps_parks.csv")
nps_parks$state <- tolower(nps_parks$state) #when I manually entered the data, the states are lower case.
us_states <- map_data("state")
map2 <- us_states |> 
  left_join(nps_parks, by = c("region" = "state")) #joins the parks data with the mapping data
ggplot(map2, aes(long, lat, group = group, fill = factor(park))) + #treated as a categorical (factor) variable rather than a continuous one
  geom_polygon(color = "black") +
  scale_fill_manual(
    values = c("1" = "darkgreen", "0" = "chocolate4"), #the data is 0 = No, 1 = yes, there is a park in this state
    name = "Has a National Park",
    labels = c("No", "Yes")
  ) +
  theme_void() +
  labs(title = "US States with a National Park",
       caption = "Data Source: National Park Service & Wikipedia (Manual Entry)")

```




This choropleth map of the United States showing which states have at least one national park. States with at least one national park are shaded in dark green, while those without any are shaded in brown (the two colors of the National Parks Service!). The maps shows that the majority of US States has at least one national park in their state. The data on whether or not the states have a national park comes from the National Park Service and Wikipedia, then manually entered into a csv file. 


**INTERACTIVE MAPS: LINK IS BELOW**

https://github.com/ParkerMax/SDS264_F24/blob/main/Parker_Max_Mini_Project_1.qmd

